Cover Your Basis: Comprehensive Data-Driven Characterization of the Binary Black Hole Population

### 2022-10-11 src/tex/methods.tex
fix up some results/methods a bit

For the method of our data-driven model, we look to the use of basis splines, or B-Splines \citep{deBoor78}. B-Splines of order $k$, are a set of order $k-1$ polynomials that 
span the space of possible spline functions interpolated from a given set of knot locations. The basis representation of the splines allows for the computational 
expensive interpolation to be just a single preprocessing step -- amortizing the cost of model evaluation during inference. To mitigate the unwanted side effects of 
adding extra knots and to avoid running a model grid of differing numbers of knots (as in \citet{Edelman_2022ApJ}), we use the smoothing prior for Bayesian P-Splines, 
allowing the data to pick the optimal scale needed to fit the present features \citep{eilers2021practical,BayesianPSplines,Jullion2007RobustSO}. We discuss basis splines, the smoothing prior 
and our specific prior choices on hyperparameters in Appendix \ref{sec:basis_splines}, \ref{sec:psplines} and \ref{sec:modelpriors}.

We parameterize each binaries' masses with the primary (more massive component) mass 
and the mass ratio ($q=\frac{m_2}{m_1}$), bounded in the range from 0 to 1. Furthermore, we model 4 of the 6 total 
spin degrees of freedom of a binary merger with each component spin magnitude $a_1$ and $a_2$, along with the cosine of the tilt angle of each component, 
$\cos{\theta_1}$ and $\cos{\theta_2}$. The tilt angle is defined as the angle between each components' spin vector and the binaries orbital angular momentum vector. 
We assume the polar spin angles are uniformly distributed in the orbital plane. For the primary mass distribution, we model the log probability with a B-Spline interpolated over 
knots linearly spaced in $\log(m_1)$ in the range of support given our observations ($[6.5\msun -- 100\msun]$), such that $\log(p(m_1 | \bm{c})) \propto B_{k}(\log(m_1) | \bm{c})$. 
We follow the same procedure for the models in mass ratio and spin distributions with knots spaced linearly across each domain, so that we have $\log(p(\theta | \bm{c}_\theta)) \propto B_{k}(\theta | \bm{c}_\theta)$, 
where $\theta$ can be $q$, $a_1$, $a_2$, $\cos{\theta_1}$ or $\cos{\theta_2}$. For the spin magnitude and tilt distributions we construct two versions of the model: first, we model  
each component's distribution is independently and identically distribution (IID), where we have a single B-Spline model and parameters (coefficients) for each binary spin. 
Secondly, we model each component's distribution to be unique, fitting separate sets of coefficients for the B-Spline models of the primary and secondary spin distributions. 
Lastly, we fit a population model on the redshift or luminosity distance distribution of BBHs, assuming a $\Lambda\mathrm{CDM}$ cosmology defined by the parameters 
from the Planck 2015 results \citep{Planck2015}. This defines an analytical mapping between each event's inferred luminosity distance, and it's redshift, which we now use interchangeably. 
We take a semi-parametric approach to model the evolution of the merger rate with redshift, following \citet{Edelman_2022ApJ}, that parameterizes modulations to an underlying model 
with splines (in our case basis Splines). We use the \textsc{PowerlawRedshift} model as the underlying distribution to modulate, which has a single hyperparameter, $\lambda_z$, and 
probability density defined as: $p(z|\lambda_z)\propto \frac{dV_c}{dz}(1+z)^{\lambda_z-1}$ \citep{Fishbach_2018redshift}. For more detailed descriptions of each model and 
specific prior choices used for the hyperparameters fit in Appendix \ref{sec:modelpriors}. Now that we have our comprehensive data-driven population model over 
each of the 7 modeled parameters built, we now apply it using the usual hierarchical Bayesian inference framework (see appendix \ref{sec:hierarchical_inference} for a review) in order to characterize the population of 
BBHs in the most recent catalog of GW observations, GWTC-3 \citep{GWTC3,o3b_astro_dist}.

>>>>

We construct our data-driven model with the application of basis splines, or B-Splines \citep{deBoor78}. B-Splines of order $k$, are a set of order $k-1$ polynomials that 
span the space of possible spline functions interpolated from a given set of knot locations. The basis representation of the splines allows for the computational 
expensive interpolation to be done in a single preprocessing step -- amortizing the cost of model evaluation during inference. To mitigate the unwanted side effects of 
adding extra knots and to avoid running a model grid of differing numbers of knots (as in \citet{Edelman_2022ApJ}), we use the smoothing prior for Bayesian P-Splines, 
allowing the data to pick the optimal scale needed to fit the present features \citep{eilers2021practical,BayesianPSplines,Jullion2007RobustSO}. We discuss basis splines, the smoothing prior 
and our specific prior choices on hyperparameters in Appendix \ref{sec:basis_splines}, \ref{sec:psplines} and \ref{sec:modelpriors}.

We parameterize each binaries' masses with the primary (more massive component) mass 
and the mass ratio ($q=\frac{m_2}{m_1}$), bounded in the range from 0 to 1. Furthermore, we model 4 of the 6 total 
spin degrees of freedom of a binary merger with each component spin magnitude $a_1$ and $a_2$, along with the cosine of the tilt angle of each component, 
$\cos{\theta_1}$ and $\cos{\theta_2}$. The tilt angle is defined as the angle between each components' spin vector and the binaries orbital angular momentum vector. 
We assume the polar spin angles are uniformly distributed in the orbital plane. For the primary mass distribution, we model the log probability with a B-Spline interpolated over 
knots linearly spaced in $\log(m_1)$ in the range of support given our observations ($[6.5\msun-100\msun]$), such that $\log(p(m_1 | \bm{c})) \propto B_{k}(\log(m_1) | \bm{c})$, where $B_{k}$ is the 
$k$-th order B-Spline function with basis coefficients, $\bm{c}$. We follow the same procedure for the models in mass ratio and spin distributions with knots spaced linearly across each domain, 
so that we have $\log(p(\theta | \bm{c}_\theta)) \propto B_{k}(\theta | \bm{c}_\theta)$, where $\theta$ can be $q$, $a_1$, $a_2$, $\cos{\theta_1}$ or $\cos{\theta_2}$. 
For the spin magnitude and tilt distributions we construct two versions of the model: first, we model  
each component's distribution is independently and identically distribution (IID), where we have a single B-Spline model and parameters (coefficients) for each binary spin. 
Secondly, we model each component's distribution to be unique, fitting separate sets of coefficients for the B-Spline models of the primary and secondary spin distributions. 
Lastly, we fit a population model on the redshift or luminosity distance distribution of BBHs, assuming a $\Lambda\mathrm{CDM}$ cosmology defined by the parameters 
from the Planck 2015 results \citep{Planck2015}. This defines an analytical mapping between each event's inferred luminosity distance, and it's redshift, which we now use interchangeably. 
We take a semi-parametric approach to model the evolution of the merger rate with redshift, following \citet{Edelman_2022ApJ}, that parameterizes modulations to an underlying model 
with splines (in our case basis Splines). We use the \textsc{PowerlawRedshift} model as the underlying distribution to modulate, which has a single hyperparameter, $\lambda_z$, and 
probability density defined as: $p(z|\lambda_z)\propto \frac{dV_c}{dz}(1+z)^{\lambda_z-1}$ \citep{Fishbach_2018redshift}. For more detailed descriptions of each model and 
specific prior choices used for the hyperparameters fit in Appendix \ref{sec:modelpriors}. Now that we have our comprehensive data-driven population model over 
each of the 7 modeled parameters built, we fit our model using the usual hierarchical Bayesian inference framework (see appendix \ref{sec:hierarchical_inference} for a review) in order to characterize 
the population of BBHs in the most recent catalog of GW observations, GWTC-3 \citep{GWTC3,o3b_astro_dist}.

### 2022-10-11 src/tex/results.tex
fix up some results/methods a bit

Figure \ref{fig:mass_distribution} shows the primary mass distribution inferred with our B-Spline model (red), where we see that it recovered consistent 
features with what was inferred with the \textsc{PowerlawPeak} and \textsc{PowerlawSpline} mass models \citep{Talbot_2018,o3a_pop,Edelman_2022ApJ,o3b_astro_dist}. 
In particular our B-Spline model finds a peak in merger rate density as a function of primary mass, at both $\sim10\msun$ and $\sim35\msun$ agreeing with those 
reported using the same dataset in \citet{o3b_astro_dist}. The B-Spline model finds the same feature at $\sim18\msun$ as the \textsc{PowerlawSpline} model, but also 
agreeing with the \textsc{PowerlawSpline} model, the mass distribution is more uncertain and less constrained in this region than other two confident peaks found.
Additionally, there is some disagreement at low masses where the models based on power laws show a much higher rate below $\sim8--9\msun$. This is partly due to  
the minimum mass hyperparameter, which is where the power law ``begins'' in those models. They enforce this hyperparameter as the minimum allowable primary and secondary 
masses of the catalog. This leads to inferences of $m_\mathrm{min}$ below the minimum observed primary mass in the catalog, $\sim 6.4\msun$, to account for secondary BBH masses 
that can possibly get to as low as $\sim 3\msun$. Since non-parametric models cannot provide meaningful insights where there is no data - we enforce the minimum primary mass of 
our B-Spline model to be $6.5\msun$, while secondary BBH masses are enforced to be greater than $3\msun$. The lack of observations in the low primary mass region could 
point towards signs of a decrease in merger rate from a purported ``lower mass gap'' between neutron star and BH masses, or could be a feature of selection effects which make it 
more unlikely to observe BBHs in that region of parameter space due to small sensitive volumes probed. We also find no evidence for a sharp fall off in merger rate either following 
the pileup at $\sim35\,\msun$, which is expected if such a pileup was due to pulsational pair instability supernovae (PPISNe), or where the maximum mass truncation of the power law models 
is inferred. The lack of any high mass truncation along with the most significant peak found to be at $\sim35\msun$, significantly lower than expected, may pose challenges for conventional stellar 
evolution theory. This could be a sign of the significant presence of subpopulations that may avoid PISN as the binary forms, but the confirmation of their existence cannot be determined 
with current catalog sizes.

@@

We report the B-Spline model's inferred marginal mass ratio distribution in figure \ref{fig:q_distribution}, where we can begin to see some departure from a simple 
power law. Specifically there looks to be a small decrease or flattening of the merger rate near equal mass ratios. Then we see a small decline to flat plateau from 
$q\sim0.3$ to $q\sim0.7$. The last step down in rate below $q\sim0.3$ is due to the minimum mass ratio truncation defined by: $q_\mathrm{min}=\frac{m_\mathrm{min}}{m_1}$. 
When marginalizing over the primary mass distribution with a strong peak at $10\msun$, we get a strong truncation at $q\sim0.3$ which is the minimum secondary mass, $3\msun$, 
over the most common primary mass, $10\msun$.

@@

The \textsc{Default} spin model describes the spin magnitude of both components as identical and independently distributed (IID) non-singular Beta distributions. \bruce{cite Wysocki/Talbot}
The Beta distribution provides a simple 2-parameter model that can produce a wide range of functional forms on the unit interval. However, the constraint that keeps 
the Beta distribution non-singular (i.e. $\alpha>1$ and $\beta>1$) enforces a spin magnitude that always has $p(a_i=0) = 0$. In reality, it is expected that 
all physical bodies must have at least some non-zero angular momentum, studies have proposed the existence of a distinct subpopulation of non-spinning or 
negligibly spinning black holes that can elude discovery with such a model \citep{FullerMa2019,RouletGWTC2Pop,BuildBetterSpinModels,Callister_NoEvidence,GWTC3MonashSpin}.

@@

We model the spin magnitude distributions as IID B-Spline distributions as a flexible check
to the assumptions of the \textsc{Default} model. Figure \ref{fig:iid_spinmag_dist} shows the inferred spin magnitude distribution with the B-Spline model, 
compared with the \textsc{Default} model. The B-Spline model finds broad agreement with the beta distribution, both peaking near ~0.2, with 90\% 
of BBH spins below \result{$\CIPlusMinus{\macros[BSplineIIDCompSpins][a_90percentile]}$} at 90\% credibility. The regions of parameter space where they don't agree as well are at 
the boundaries. The B-Spline model has no enforcement for vanishing support at the extremal values like the Beta distribution, 
allowing it to probe the zero-spin question. We find broad support, with large variance, for non-zero probabilities at $a_i=0$, but cannot confidently determine the presence of 
a significant non-spinning subpopulation, corroborating similar recent conclusions \citep{BuildBetterSpinModels,Callister_NoEvidence,GWTC3MonashSpin}. 
We now repeat the same analysis as above but with independent B-Spline distributions for each spin magnitude component. In figure \ref{fig:ind_spinmag_dist} 
we show the inferred primary (orange), and secondary (olive) spin magnitude distributions inferred when relaxing the IID assumption. We find both 
distributions to be remarkably similar, showing no signs that each component spin magnitude are distributed differently from each other.

@@

The \textsc{Default} spin model assumes the spin orientation of both components are identical and independently distributed (IID) mixture models over an
aligned and an isotropic component. The aligned component is modeled with a truncated Gaussian distribution with mean at $\cos{\theta}=1$ and variance a free 
hyperparameter to be fit. This provides a simple 2-parameter model motivated by simple distributions expected from the two main formation scenario families, allowing 
for a straightforward interpretation of results. One possible limitation however, is that by construction this distribution is forced to peak at perfectly aligned spins, 
i.e. $\cos{\theta}=1$. While this may be a reasonable assumption, \citet{spinitasyoulike} recently extended the model space of parametric descriptions 
used to model the spin orientation distribution and found considerable evidence that the distribution peaks away from $\cos{\theta}=1$. Again, this provides a clear 
use-case where data-driven models can help in our understanding and possibly resolve disputes among differing parametric approaches.

@@

We repeat the tests done above on the spin magnitude distributions now with the spin orientation distributions. Figure \ref{fig:iid_spintilt_dist} 
shows the inferred spin magnitude distribution with the B-Spline model, compared with the \textsc{Default} model from \citet{o3b_astro_dist}. 
The B-Spline inferences have large uncertainties but start to show the same features as found and discussed in \citet{spinitasyoulike}. 
We find a distribution that instead of intrinsically peaking at $\cos{\theta}=1$, is found to peak at: $\cos{\theta}=$\result{$\CIPlusMinus{\macros[BSplineIIDCompSpins][peakCosTilt]}$}, at 
90\% credibility. We find less, but still considerable, support for misaligned spins (i.e. $\cos{\theta}<0$) as the \textsc{Default} model. Specifically we 
find that the fraction of misaligned systems is $f_{\cos{\theta}<0}=$\result{$\CIPlusMinus{\macros[BSplineIIDCompSpins][negFrac]}$}. This implies 
the presence of an isotropic component as expected by dynamical formation channels, albeit less than with the \textsc{Default} model. To quantify the 
amount of isotropy in the tilt distribution we calculate $\log_{10}Y$, where $Y$ is the ratio of nearly aligned tilts to nearly anti-aligned, 
introduced in \citet{spinitasyoulike} and defined as:

@@

Now we model each of the components' orientation distributions with independent B-Spline models as done above, and show the inferred 
primary (orange), and secondary (olive) distributions in figure \ref{fig:ind_spintilt_dist}. The orientation distributions do not agree quite as well as 
the magnitude distributions did in figure \ref{fig:ind_spinmag_dist}, and the PPDs look to peak at different values, but the wide uncertainties still are 
consistent with both each other and the \textsc{Default} model's PPD. We find the two distributions to peak at: $\cos{\theta_1}=$\result{$\CIPlusMinus{\macros[BSplineIndependentCompSpins][peakCosTilt1]}$} 
and $\cos{\theta_2}=$\result{$\CIPlusMinus{\macros[BSplineIndependentCompSpins][peakCosTilt2]}$}, showing that the primary distribution peak is inferred further away 
from the assumed $\cos{\theta}=1$ with the \textsc{Default} model. There is also significant (albeit uncertain) evidence of spin misalignment in each distribution, finding 
the fraction of misaligned primary and secondary components as: $f_{\cos{\theta_1}<0}=$\result{$\CIPlusMinus{\macros[BSplineIndependentCompSpins][negFrac1]}$} and 
$f_{\cos{\theta_2}<0}=$\result{$\CIPlusMinus{\macros[BSplineIndependentCompSpins][negFrac2]}$}. We again calculate $\log_{10}Y$  
for each component distribution and find: $\log_{10}Y_1=$\result{$\CIPlusMinus{\macros[BSplineIndependentCompSpins][log10gammaFrac1]}$} and 
$\log_{10}Y_2=$\result{$\CIPlusMinus{\macros[BSplineIndependentCompSpins][log10gammaFrac2]}$}.

@@

\begin{figure}[h]
    \includegraphics[width=\linewidth]{figures/rate_vs_z_plot.pdf}
    \caption{The BBH merger rate as a function of redshift. We show the B-Spline model (red), with 16 knots spaced linearly in $\log(z)$, 
    from the minimum to the maximum observed redshifts. The solid line shows the population predictive distribution (PPD) and the shaded region, 
    the 90\% credible interval. We show the inferred PPD from the \textsc{PowerlawRedshift} (blue) model from the LVK's GWTC-3 population analyses \citep{o3b_astro_dist}.}
    \label{fig:rofz}
    \script{rate_vs_z_plot.py}
\end{figure}

>>>>

Figure \ref{fig:mass_distribution} shows the primary mass distribution inferred with our B-Spline model (red), where we see that it recovered consistent 
features with what was inferred with the \textsc{PowerlawPeak} and \textsc{PowerlawSpline} mass models \citep{Talbot_2018,o3a_pop,Edelman_2022ApJ,o3b_astro_dist}. 
In particular our B-Spline model finds a peak in merger rate density as a function of primary mass, at both $\sim10\msun$ and $\sim35\msun$ agreeing with those 
reported using the same dataset in \citet{o3b_astro_dist}. The B-Spline model finds the same feature at $\sim18\msun$ as the \textsc{PowerlawSpline} model, but also 
agreeing with the \textsc{PowerlawSpline} model, the mass distribution is more uncertain and less constrained in this region than other two confident peaks found.
Additionally, there is some disagreement at low masses where the models based on power laws show a much higher rate below $\sim8--9\msun$. This is partly due to  
the minimum mass hyperparameter, which is where the power law ``begins'' in those models. They enforce this hyperparameter as the minimum allowable primary and secondary 
masses of the catalog. This leads to inferences of $m_\mathrm{min}$ below the minimum observed primary mass in the catalog, $\sim 6.4\msun$, to account for secondary BBH masses 
that can possibly get to as low as $\sim 3\msun$. Since non-parametric models do not provide meaningful insights where there is no data - we enforce the minimum primary mass of 
our B-Spline model to be $6.5\msun$, while secondary BBH masses are enforced to be greater than $3\msun$. The lack of observations in the low primary mass region could 
point towards signs of a decrease in merger rate from a purported ``lower mass gap'' between neutron star and BH masses, or could be a feature of selection effects which make it 
more unlikely to observe BBHs in that region of parameter space due to small sensitive volumes probed. We also find no evidence for a sharp fall off in merger rate either following 
the pileup at $\sim35\,\msun$, which is expected if such a pileup was due to pulsational pair instability supernovae (PPISNe), or where the maximum mass truncation of the power law models 
is inferred. The lack of any high mass truncation along with the most significant peak found to be at $\sim35\msun$, significantly lower than expected, may pose challenges for conventional stellar 
evolution theory. This could be hinting at the significant presence of subpopulations that avoid PISN during binary formation, but the confirmation of the existence of such subpopulations 
cannot be determined with current catalog sizes.

@@

We report the B-Spline model's inferred marginal mass ratio distribution in figure \ref{fig:q_distribution}, where we can begin to see departure from a simple 
power law. Specifically there looks to be a small decrease or flattening of the merger rate near equal mass ratios, as well as a fatter tail towards unequal mass ratios 
than the power law based models find. Then we see a small decline to flat plateau from $q\sim0.3$ to $q\sim0.7$. The sharp decrease in rate just below 
$q\sim0.3$ is due to the minimum mass ratio truncation defined by: $q_\mathrm{min}=\frac{m_\mathrm{min}}{m_1}$. When marginalizing over the primary mass 
distribution with a strong peak at $10\msun$, the mass ratio distribution truncates at $q\sim0.3$: the minimum secondary mass, $3\msun$, 
divided by the most common primary mass, $\sim10\msun$.

@@

The \textsc{Default} spin model describes the spin magnitude of both components as identical and independently distributed (IID) non-singular Beta distributions \citep{Talbot_2017,Wysocki_2019}.
The Beta distribution provides a simple 2-parameter model that can produce a wide range of functional forms on the unit interval. However, the constraint that keeps 
the Beta distribution non-singular (i.e. $\alpha>1$ and $\beta>1$) enforces a spin magnitude that always has $p(a_i=0) = 0$. Recent studies have proposed the possible existence of a 
distinct subpopulation of non-spinning or negligibly spinning black holes that can elude discovery with such a model \citep{FullerMa2019,RouletGWTC2Pop,BuildBetterSpinModels,Callister_NoEvidence,GWTC3MonashSpin}.

@@

We model the spin magnitude distributions as IID B-Spline distributions as a flexible check to the assumptions of the \textsc{Default} model. 
Figure \ref{fig:iid_spinmag_dist} shows the inferred spin magnitude distribution with the B-Spline model, compared with the \textsc{Default} model. 
The B-Spline model finds consistent agreement with the beta distribution, peaking near $a\sim0.2$, with 90\% of BBH spins below 
\result{$\CIPlusMinus{\macros[BSplineIIDCompSpins][a_90percentile]}$} at 90\% credibility. The B-Spline model has no enforcement for vanishing support at the extremal values like the Beta distribution, 
allowing it to probe the zero-spin question. We find broad support, with large variance, for non-zero probabilities at $a_i=0$, but cannot confidently determine the presence of 
a significant non-spinning subpopulation, corroborating similar recent conclusions \citep{BuildBetterSpinModels,Callister_NoEvidence,GWTC3MonashSpin}. 
We now repeat the same analysis as above but with independent B-Spline distributions for each spin magnitude component. In figure \ref{fig:ind_spinmag_dist} 
we show the inferred primary (orange), and secondary (olive) spin magnitude distributions inferred when relaxing the IID assumption. We find that the primary spin magnitude distribution 
peaking slightly higher at $a\sim0.25$ than the IID B-Spline model, but with similar support at near vanishing spins. The secondary spin magnitude distribution is more uncertain due to the 
higher measurement uncertainty in inferring the secondary spins of BBH systems. The secondary distribution also peaks at smaller spin magnitudes of $a\sim0.15$, showing considerably larger 
PPD and confidence band at $a\sim0$ than the primary distribution or B-Spline IID spin magnitude distribution in figure \ref{fig:iid_spinmag_dist}. While the distributions are still broadly 
consistent with each other this could also be signs pointing towards component spin magnitude distributions uniquely distributed.

@@

The \textsc{Default} spin model also assumes the spin orientation of both components are identical and independently distributed (IID), with a mixture model over an
aligned and an isotropic component. The aligned component is modeled with a truncated Gaussian distribution with mean at $\cos{\theta}=1$ and variance a free 
hyperparameter to be fit \citep{Talbot_2017,Wysocki_2019,o3a_pop,o3b_astro_dist}. This provides a simple 2-parameter model motivated by simple distributions expected from the two main formation scenario families, allowing 
for a straightforward interpretation of results. One possible limitation however, is that by construction this distribution is forced to peak at perfectly aligned spins, 
i.e. $\cos{\theta}=1$. While this may be a reasonable assumption, \citet{spinitasyoulike} recently extended the model space of parametric descriptions 
used to model the spin orientation distribution and found considerable evidence that the distribution peaks away from $\cos{\theta}=1$. Again, this provides a clear 
use-case where data-driven models can help in our understanding of and possibly resolve disputes among differing parametric approaches to modeling the spin orientation distribution.

@@

Figure \ref{fig:iid_spintilt_dist} shows the inferred spin magnitude distribution with the IID spin B-Spline model, compared with the \textsc{Default} model from \citet{o3b_astro_dist}. 
The B-Spline inferences have large uncertainties but start to show the same features as found and discussed in \citet{spinitasyoulike}. 
We find a distribution that instead of intrinsically peaking at $\cos{\theta}=1$, is found to peak at: $\cos{\theta}=$\result{$\CIPlusMinus{\macros[BSplineIIDCompSpins][peakCosTilt]}$}, at 
90\% credibility. We find less, but still considerable, support for misaligned spins (i.e. $\cos{\theta}<0$) as the \textsc{Default} model. Specifically we 
find that the fraction of misaligned systems is $f_{\cos{\theta}<0}=$\result{$\CIPlusMinus{\macros[BSplineIIDCompSpins][negFrac]}$}. This implies 
the presence of an isotropic component as expected by dynamical formation channels, albeit less than with the \textsc{Default} model. To quantify the 
amount of isotropy in the tilt distribution we calculate $\log_{10}Y$, where $Y$ is the ratio of nearly aligned tilts to nearly anti-aligned, 
introduced in \citet{spinitasyoulike} and defined as:

@@

Now we model each of the components' orientation distributions with independent B-Spline models as done above, and show the inferred 
primary (orange), and secondary (olive) distributions in figure \ref{fig:ind_spintilt_dist}. The orientation distributions do not match perfectly, but both are broadly
are consistent with each other and the \textsc{Default} model's PPD given the wide credible intervals. We find the two distributions to peak at: $\cos{\theta_1}=$\result{$\CIPlusMinus{\macros[BSplineIndependentCompSpins][peakCosTilt1]}$} 
and $\cos{\theta_2}=$\result{$\CIPlusMinus{\macros[BSplineIndependentCompSpins][peakCosTilt2]}$}, showing that the primary distribution peak is inferred further away 
from the assumed $\cos{\theta}=1$ with the \textsc{Default} model. There is also significant (albeit uncertain) evidence of spin misalignment in each distribution, finding 
the fraction of misaligned primary and secondary components as: $f_{\cos{\theta_1}<0}=$\result{$\CIPlusMinus{\macros[BSplineIndependentCompSpins][negFrac1]}$} and 
$f_{\cos{\theta_2}<0}=$\result{$\CIPlusMinus{\macros[BSplineIndependentCompSpins][negFrac2]}$}. We again calculate $\log_{10}Y$  
for each component distribution and find: $\log_{10}Y_1=$\result{$\CIPlusMinus{\macros[BSplineIndependentCompSpins][log10gammaFrac1]}$} and 
$\log_{10}Y_2=$\result{$\CIPlusMinus{\macros[BSplineIndependentCompSpins][log10gammaFrac2]}$}.

@@

\begin{figure}[h]
    \includegraphics[width=\linewidth]{figures/rate_vs_z_plot.pdf}
    \caption{The BBH merger rate as a function of redshift. We show the B-Spline model (red), with 16 knots spaced linearly in $\log(z)$, 
    from the minimum to the maximum observed redshifts. The solid line shows the population predictive distribution (PPD) and the shaded region, 
    the 90\% credible interval. We show the inferred 90\% credible interval from the \textsc{PowerlawRedshift} model 
    from the LVK's GWTC-3 population analyses in blue and a power law with exponent of 2.7 in gray, representing the expected star formation rate 
    \citep{Madau_2014, o3b_astro_dist}.}
    \label{fig:rofz}
    \script{rate_vs_z_plot.py}
\end{figure}
